---
title: "MichiganCOVID19Stats"
author: "hollen"
date: "3/24/2020"
output:
  pdf_document: default
  html_document: default
params:
  mergeToday: 0
---

```{r setup, include=FALSE, echo = FALSE, warning = FALSE, error = FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(rvest)
library(tidyverse)
library(knitr)
```

## Michigan COVID-19 Evaluation

During the COVID-19 outbreak, during which Michiganders were asked to remain at home and "Stay Home, Stay Safe," I decided to spend a bit of time to collect the outbreak reports from http://michigan.gov/coronavirus so that I could track the outbreak of the virus from each day's report.  

To date I have not yet found a repository of all of the data collected so far in the state of Michigan, so I will use this application to collect the results from each day and merge them together in an accumulated data set, which I can store as an RData file as well as export to CSV for sharing with other Michiganders.  (in a future release!)

#### Website Cummulative Totals for the Day

```{r WebSiteCummTotals, echo = FALSE, warning = FALSE, error = FALSE}
uriMiCummTots <- "https://www.michigan.gov/coronavirus/0,9753,7-406-98163-520743--,00.html"
txtMiCummTots <- read_html(uriMiCummTots)
tblMiCummTots <- html_table(txtMiCummTots, header = FALSE, fill = TRUE)
dfMiCummTots <- as_tibble(tblMiCummTots[[1]])
dfMiCummTots <- dfMiCummTots %>%
  transform(County = X1, 
            TotalCases = as.numeric(X2), 
            TotalDeaths = as.numeric(X3)) %>%
  select(-X1, -X2, -X3) %>%
  filter(County != 'County') %>%
  filter(County != 'Total') %>% 
  mutate(County = as.character(County)) %>%
  arrange(County)
kable(dfMiCummTots)
```

This reports a total of `r dfMiCummTots %>% select(TotalCases) %>% sum(na.rm = TRUE)` cases with `r dfMiCummTots %>% select(TotalDeaths) %>% sum(na.rm = TRUE)` deaths.

```{r ScrapeMichiganDailyValues, echo = FALSE, warning = FALSE, error = FALSE}
uriMichiganDaily <- "https://www.michigan.gov/coronavirus"
txtMichiganDaily <- read_html(uriMichiganDaily)
tblMichiganDaily <- html_table(txtMichiganDaily, header = FALSE, fill = TRUE)
dfMichiganDaily <- as_tibble(tblMichiganDaily[[1]])
dfMichiganDaily <- dfMichiganDaily %>%
  transform(County = X1, 
            TotalCases = as.numeric(X2), 
            TotalDeaths = as.numeric(X3)) %>%
  select(-X1, -X2, -X3) %>%
  filter(County != 'County') %>%
  filter(County != 'Total') %>% 
  mutate(Date = Sys.Date()) %>%
  mutate(County = as.character(County)) %>%
  arrange(desc(TotalCases))
```

```{r PersistWithAccumulatedValues, echo = FALSE, warning = FALSE, error = FALSE}
load(file='dfCOVID19MichiganTotals.Rdata')
# Append the new values to the current accumulated values
if (params$mergeToday > 0) {
  dfCOVID19MichiganTotals <- rbind(dfCOVID19MichiganTotals, dfMichiganDaily)
  save(dfCOVID19MichiganTotals, file='dfCOVID19MichiganTotals.Rdata')
  write_csv(dfCOVID19MichiganTotals, path='michigan_covid19_cases.csv')
}
dfCOVID19MichiganTotals <- dfCOVID19MichiganTotals %>%
  mutate(County = as.character(County)) %>%
  arrange(County)
```

#### My Cummulative Totals to Date

These are based on previously reported daily values which I'm collecting in this dataset and report as well as cases reported today:

```{r DetermineMyAccumulatedTotals, echo = FALSE, warning = FALSE, error = FALSE}
dfResult <- dfCOVID19MichiganTotals %>%
  group_by(County) %>%
  summarise(AllCases=sum(TotalCases, na.rm=TRUE), AllDeaths=sum(TotalDeaths, na.rm=TRUE)) %>%
  arrange(County)
kable(dfResult)
```

My data reflects a total of `r dfCOVID19MichiganTotals %>% select(TotalCases) %>% sum(na.rm = TRUE)` cases with `r dfCOVID19MichiganTotals %>% select(TotalDeaths) %>% sum(na.rm = TRUE)` deaths.

#### Top Ten Cases by County from my Cummulative Data

```{r TopTenCasesByArea, echo = FALSE, warning = FALSE, error = FALSE}
dfResult <- dfCOVID19MichiganTotals %>% 
  group_by(County) %>% 
  summarise(AllCases=sum(TotalCases, na.rm = TRUE)) %>% 
  arrange(desc(AllCases)) %>%
  top_n(10)
kable(dfResult)
```

#### Top Ten Deaths by County from my Cummulative Data

```{r TopTenDeathsByArea, echo = FALSE, warning = FALSE, error = FALSE}
dfResult <- dfCOVID19MichiganTotals %>% 
  group_by(County) %>% 
  summarise(AllDeaths=sum(TotalDeaths, na.rm = TRUE)) %>%
  filter(AllDeaths > 0) %>%
  arrange(desc(AllDeaths)) %>%
  top_n(10)
kable(dfResult)
```



